---
title: "R Assignment 1 - Summer"
author: "Sanmesh Sanjay Shintre"
format:
  beamer:
    theme: Madrid
    colortheme: seahorse
    fontsize: 8pt
    toc: true
    slide-level: 2
    keep-tex: true
    includes:
      in-header: custom.tex
editor: visual
---

# Use data.table to read in the data

```{r, echo=TRUE, results='show'}
library(data.table)
library(ggplot2)  
library(knitr)
library(dplyr)

#Loading the datasets 
dt_india <- fread("indicators_ind.csv")[-1,]
dt_are <- fread("indicators_are.csv")[-1,]
dt_usa <- fread("indicators_usa.csv")[-1,]
```

-   Loading the packages required.

-   Loaded the dataset using fread for 3 countries.

# Assign the correct class to the variables

```{r, echo=TRUE, results='show'}
options(width = 70)
dt_india[, `:=`(
  `Country Name` = as.factor(`Country Name`),
  `Country ISO3` = as.factor(`Country ISO3`),
  `Indicator Name` = as.factor(`Indicator Name`),
  `Indicator Code` = as.factor(`Indicator Code`),
  Year = as.integer(Year),
  Value = as.numeric(Value)
)]
dt_are[, `:=`(
  `Country Name` = as.factor(`Country Name`),
  `Country ISO3` = as.factor(`Country ISO3`),
  `Indicator Name` = as.factor(`Indicator Name`),
  `Indicator Code` = as.factor(`Indicator Code`),
  Year = as.integer(Year),
  Value = as.numeric(Value)
)]
dt_usa[, `:=`(
  `Country Name` = as.factor(`Country Name`),
  `Country ISO3` = as.factor(`Country ISO3`),
  `Indicator Name` = as.factor(`Indicator Name`),
  `Indicator Code` = as.factor(`Indicator Code`),
  Year = as.integer(Year),
  Value = as.numeric(Value)
)]

```

## Combining 3 datasets

```{r, echo=TRUE, results='show'}
dt_all <- rbindlist(list(dt_india, dt_are, dt_usa)
                    , use.names = TRUE, fill = TRUE)
```

-   Assigned the correct class for each of the column of dataset for respective countries.
-   Such as integer and numeric for Year and Value respectively.
-   Combined 3 datasets into one single dataset using rbindlist().

# Data Exploration

```{r, echo=TRUE, results='show'}
dt_all[, .N, by = .(`Country Name`, 
                    `Indicator Name`)][, .N, by = `Country Name`]
common_indicators <- Reduce(intersect, list(
  unique(dt_india$`Indicator Name`),
  unique(dt_are$`Indicator Name`),
  unique(dt_usa$`Indicator Name`)
))
length(common_indicators)  
```

-   Taking the common indicators from dt_all to common_indicators.

-   Printing the number of common_indicators in the dataset.

# Data Exploration and plots

```{r, echo=TRUE, results='show'}
indicator_counts <- dt_all[, .N, by = .(`Country Name`, `Indicator Name`)][, .N, by = `Country Name`]

top_indicators_overall <- dt_all %>%
  group_by(`Indicator Name`) %>%
  summarise(Count = n(), .groups = "drop") %>%
  slice_max(order_by = Count, n = 15)
```

-   Count **how many indicators** each country has: two‐step grouping with .N.

-   Prepare two exploratory outputs:

    1.  **indicator_counts** (per country)

    2.  **top_indicators_overall** (top 15 by frequency)

# Data Analysis using data.table (with keyby)

```{r, echo=TRUE, results='show'}
dt_filtered <- dt_all[`Indicator Name` %in% common_indicators]
dt_migration_summary <- dt_filtered[
  `Indicator Name` == "Net migration",
  .(Average_Migration = mean(Value, na.rm = TRUE)),
  keyby = .(Year, `Country Name`)
]
dt_fuel_exports <- dt_filtered[`Indicator Name` ==
                                 "Fuel exports (% of merchandise exports)"]
dt_fuel_summary <- dt_fuel_exports[,.(
    Avg_Export = mean(Value, na.rm = TRUE),
    SD_Export = sd(Value, na.rm = TRUE),
    Min = min(Value, na.rm = TRUE),
    Max = max(Value, na.rm = TRUE),
    Observations = .N
  ),keyby = `Country Name`
]


```

## Data Analysis using data.table (with keyby)

```{r, echo=TRUE, results='show'}
dt_fuel_summary
```

-   **Filter** to only the common_indicators.

-   Use **keyby** to efficiently compute:

    -   **Average Net Migration** by Year & Country
    -   **Fuel Exports** summary stats by Country

# Plots

## Plot 1 of Analysis

```{r, echo=TRUE, results='show',fig.width=6, fig.height=2.4}
ggplot(indicator_counts, aes(x = `Country Name`, y = N, fill = `Country Name`)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    title = "Number of Unique Indicators per Country",
    x = "Country",
    y = "Number of Indicators"
  ) +
  theme(legend.position = "none")


```

## Plot 2 of Analysis

```{r, echo=TRUE, results='show',fig.width=6, fig.height=2.4}
ggplot(top_indicators_overall, aes(x = reorder(`Indicator Name`, Count), y = Count)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 15 Most Common Indicators in Dataset",
    x = "Indicator",
    y = "Observation Count"
  ) +
  theme_minimal(base_size = 8)

```

## Plot 3 of Analysis

```{r, echo=TRUE, results='show',fig.width=6, fig.height=2.4}
ggplot(dt_migration_summary, aes(x = Year, y = Average_Migration, 
                                 color = `Country Name`)) +
  geom_line(linewidth = 1.2) +
  theme_minimal() +
  labs(
    title = "Average Net Migration Over Time",
    x = "Year",
    y = "Average Migration",
    color = "Country"
  )
```

## Plot 4 of Analysis

```{r, echo=TRUE, results='show',fig.width=6, fig.height=2.4}
ggplot(dt_fuel_summary, aes(x = `Country Name`, y = Avg_Export, 
                            fill = `Country Name`)) +
  geom_col(width = 0.6) +
  theme_minimal() +
  labs(
    title = "Average Fuel Exports (% of Merchandise Exports)",
    x = "Country", y = "Average % of Fuel Exports"
  ) +
  theme(legend.position = "none")
```
