---
title: "Quarto Assignment"
author: "Sanmesh Shintre (24219242)"
format: pdf
execute:
  echo: false 
  warning: false
  message: false
---

# *Introduction*


The **`ToothGrowth`** dataset is available in base R and contains data on the effect of **vitamin C** on tooth growth in guinea pigs.

There are **60 observations** and **3 variables**: - `len`: Tooth length - `supp`: Supplement type (VC = ascorbic acid, OJ = orange juice) - `dose`: Dose in milligrams/day (0.5, 1.0, or 2.0)

This document explores how tooth length changes with different **vitamin C doses**, grouped by supplement type.

------------------------------------------------------------------------

# *Descriptive Statistics*

```{r}
# Summarizing the tooth length by dose and supplement type
aggregate(len ~ dose + supp, data = ToothGrowth, FUN = mean)

```

We calculate the **average tooth length** for each combination of dose and supplement type.

# *Boxplot: Tooth Length by Dose and Supplement*

```{r}
# Creating a grouped boxplot using the ggplot2
library(ggplot2)

ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = supp)) +
  geom_boxplot() +
  labs(title = "Tooth Length by Vitamin C Dose and Supplement Type",
       x = "Dose (mg/day)",
       y = "Tooth Length") +
  scale_fill_manual(values = c("orange", "steelblue")) +
  theme_minimal()
```

*The boxplot shows that tooth length increases with higher doses of vitamin C.* Orange juice (`OJ`) seems slightly more effective at lower doses compared to ascorbic acid (`VC`).

# *Conclusion*

From the summary and plot, we observe:

1)  Tooth length generally increases with **higher doses** of vitamin C.

2)  Supplement type also appears to influence growth, especially at **lower doses**.

3)   The `ToothGrowth` dataset provides a great example of visualizing experimental results using R.
